### 2015-09-24 presentazione/presentazione.tex
stampa

\documentclass[11pt]{beamer}
%\usetheme{Montpellier}
\usetheme{Dresden}
\usepackage[utf8]{inputenc}
\usepackage[italian]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[compatibility=false]{caption}
\usepackage{subcaption}
\usepackage{multicol}
\author{Davide Bazzanella}
\title{Thermal tuning of phase-matching in a multimodal SOI waveguide with $\chi^{(3)}$ non linearity}
\date{28 settembre 2015} 
%\textsc{\LARGE \textbf{DEPARTMENT OF PHYSICS}}\\[0.5cm]
%\textsc{\LARGE \textbf{DEGREE IN PHYSICS – LAUREA IN FISICA}}\\[2.5cm]
%\textsc{\Large THESIS}\\[1.3cm]

%\setbeamercovered{transparent} 
\setbeamertemplate{navigation symbols}{} 
%\logo{} 
%\institute{} 
%\subject{}

\usepackage{lmodern}
\newcommand{\frameofframes}{/}
\newcommand{\setframeofframes}[1]{\renewcommand{\frameofframes}{#1}}

\setframeofframes{of}
\makeatletter
\setbeamertemplate{footline}
  {%
    \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
    \end{beamercolorbox}
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
      leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
      \leavevmode{\usebeamerfont{author in head/foot}\insertshortauthor}%
      \hfill%
      {\usebeamerfont{institute in head/foot}\usebeamercolor[fg]{institute in head/foot}\insertshortinstitute}%
    \end{beamercolorbox}%
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
      leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
      {\usebeamerfont{title in head/foot}\insertshorttitle}%
      \hfill%
      {\usebeamerfont{frame number}\usebeamercolor[fg]{frame number}\insertframenumber~\frameofframes~\inserttotalframenumber}
    \end{beamercolorbox}%
    \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
    \end{beamercolorbox}
  }
\makeatother

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}{Outline}
\footnotesize \tableofcontents
\end{frame}

\section{Introduction}
ciao
\section{Nonlinear Optics}
\subsection{Four-Wave Mixing}
\subsection{Conservation of energy and momentum}
\subsubsection{Multimodal Waveguides}
\subsection{Peak tuning}
\subsubsection{Thermo-Optic Coefficient}
\section{Computational model}
\subsection{Mode simulation}
\subsection{•}
\subsection{Results}
\section{Conclusion}

\end{document}