### 2015-09-26 presentazione/presentazione.tex
final version?

\documentclass[11pt, leqno]{beamer}
%\usetheme{Montpellier}
\usetheme{Dresden}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[compatibility=false]{caption}
\usepackage{subcaption}
\usepackage{multicol}

\usepackage{siunitx}
\usepackage{wrapfig}
%\usepackage{float}

\usepackage{caption}
\usepackage{subcaption}

\author{Davide Bazzanella}
\title{Thermal tuning of phase-matching in a multimodal SOI waveguide with $\chi^{(3)}$ non linearity}
\date{28 settembre 2015} 
%\textsc{\LARGE \textbf{DEPARTMENT OF PHYSICS}}\\[0.5cm]
%\textsc{\LARGE \textbf{DEGREE IN PHYSICS – LAUREA IN FISICA}}\\[2.5cm]
%\textsc{\Large THESIS}\\[1.3cm]

%\setbeamercovered{transparent} 
\setbeamertemplate{navigation symbols}{} 
%\logo{} 
%\institute{} 
%\subject{}

\usepackage{lmodern}
\newcommand{\frameofframes}{/}
\newcommand{\setframeofframes}[1]{\renewcommand{\frameofframes}{#1}}

\setframeofframes{of}
\makeatletter
\setbeamertemplate{footline}
  {%
    \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
    \end{beamercolorbox}
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
      leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
      \leavevmode{\usebeamerfont{author in head/foot}\insertshortauthor}%
      \hfill%
      {\usebeamerfont{institute in head/foot}\usebeamercolor[fg]{institute in head/foot}\insertshortinstitute}%
    \end{beamercolorbox}%
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
      leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
      {\usebeamerfont{title in head/foot}\insertshorttitle}%
      \hfill%
      {\usebeamerfont{frame number}\usebeamercolor[fg]{frame number}\insertframenumber~\frameofframes~\inserttotalframenumber}
    \end{beamercolorbox}%
    \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
    \end{beamercolorbox}
  }
\makeatother

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}{Outline}
%\footnotesize
\tableofcontents
\end{frame}

\section{Introduction}
\begin{frame}{Introduction}

	\textbf{Research field}:\\
	bio-sensors, gas sensors and bio-medical instrumentation

	\vspace{5pt}
	\textbf{Problem}:\\
	lack of sources/sensors in the MIR (3-10\si{\um})

	\begin{figure}
		\centering
		\includegraphics[width=\textwidth]{spectrum2.png}
	\end{figure}
\end{frame}
\begin{frame}
	\textbf{Solution}:\\
	frequency down convertion from NIR/VIS sources to MIR\\
	frequency up convertion from MIR signals to NIR/VIS sensor

	\begin{figure}[hb]
%		\centering
%		\begin{subfigure}[c]{.725\textwidth}
			\centering
			\includegraphics[width=.75\textwidth]{upconversion2_short.png}
%			\caption{}
%		\end{subfigure}
%		\hspace{0.02\textwidth}
%		\begin{subfigure}[c]{.225\textwidth}
%			\centering
%			\includegraphics[width=\textwidth]{upconverter_HITACHI.png}
%			\caption{}
%		\end{subfigure}
	\end{figure}	
%	\vspace{15pt}
	\textbf{Aim}:\\
	Develop integrated frequency converter tunable with temperature
\end{frame}

\section{Nonlinear Optics}
\begin{frame}{Nonlinear Optics}
	Frequency convertion employs nonlinear properties of materials
	\begin{equation}
		\textbf{P} = \varepsilon_0 \left( \chi^{(1)} \cdot \textbf{E} + \chi^{(2)} : \textbf{E}^2 + \chi^{(3)} \vdots \textbf{E}^3 + \cdots \right) = \mathrm{\textbf{P}_L} + \mathrm{\textbf{P}_{NL}}
		\label{eq_P_general}
	\end{equation}
	\begin{wrapfigure}{r}{0.48\textwidth}
		\centering
		\includegraphics[width=.5\textwidth]{ppln-ovens2.jpg}
	\end{wrapfigure}
	\begin{itemize}
	\item	$\chi^{(1)}$	:	linear part
	\item	$\chi^{(2)}$	:	2$^{nd}$ order nonlinearity\\ \hspace{25pt} (SHG, TWM)
	\item	$\chi^{(3)}$	:	3$^{rd}$ order nonlinearity\\ \hspace{25pt} (THG, FWM)
	\end{itemize}
\end{frame}
%\subsection{Four-Wave Mixing}
\begin{frame}{Four-Wave Mixing}
	\begin{wrapfigure}{r}{0.33\textwidth}
		\vspace{-20pt}
	    \includegraphics[width=.33\textwidth]{photons.png}
	\end{wrapfigure}
%	\begin{wrapfigure}{r}{0.35\textwidth}
%		\vspace{-20pt}
%	    \includegraphics[width=.33\textwidth]{W&K.png}
%	\end{wrapfigure}
	FWM uses 3$^{rd}$ order nonlinearity $\chi^{(3)}$.
	%\begin{wrapfigure}{r}{0.22\textwidth}
	%    \includegraphics[width=.22\textwidth]{momentum.png}
	%\end{wrapfigure}
	\begin{wrapfigure}{r}{0.45\textwidth}
		\vspace{-55pt}
	    \includegraphics[width=.45\textwidth]{frequency.png}
	\end{wrapfigure}

	\vspace{10pt}
	It consists of the interaction between four EM waves of different frequencies.
	\begin{align}
	\omega_1 + \omega_2 &= \omega_3 + \omega_4\\
	k_1 + k_2 &= k_3 + k_4
	\label{eq_general_f}
	\end{align}

	\vspace{5pt}
	Special case:\\ $\omega_1 = \omega_2 = \omega_P$, with stimulation.
\end{frame}
%\subsection{Conservation of energy and momentum}
\begin{frame}{Conservation of energy and momentum}
	\begin{figure}
		\centering
		\vspace{10pt}
		\includegraphics[width=.8\textwidth]{setup_b&w.png}
	\end{figure}
	Conservation of energy is naturally verified
	\begin{equation}
		\omega_I = 2\omega_P - \omega_S = \omega_P \pm \Omega
		\label{eq_energy}
	\end{equation}
	Conservation of momentum is imposed in the phase-mismatch
	\begin{equation}
		\Delta k = 2\,k_P(\omega_P) - k_S(\omega_S) -k_I(\omega_I) = 0
		\label{eq_momentum}
	\end{equation}
	%\begin{equation*}
	%\mathrm{where} \qquad k = \frac{\omega}{c} = \frac{2\pi}{\lambda}
	%\end{equation*}
\end{frame}
\begin{frame}
	Generation happens also for $\Delta k \approx 0$, but at lower efficiency.
	\begin{wrapfigure}{r}{0.5\textwidth}
		\vspace{10pt}
		\centering
	    \includegraphics[width=.5\textwidth]{sinc2.png}
	\end{wrapfigure}
	\begin{equation}
	\begin{split}
		I_I &\propto \left| \int_0^L exp(\mathrm{i}\Delta k z)dz\right| \\
			&\propto L^2 sinc^2(\Delta k L/ 2\pi)
	\end{split}
	\end{equation}
	We define the coherence length
	\begin{equation}
		L_{coh} = \frac{2\pi}{|\Delta k|}
	\label{eq_coherence_length}
	\end{equation}
\end{frame}
\begin{frame}
	Bulk silicon does not permit to verify both equations.
	\begin{equation}
	k(\omega) = k_0 n(\omega)
	\label{eq_k(w)}
	\end{equation}
	\begin{figure}
		\centering
%		\vspace{-5pt}
		\includegraphics[width=0.8\textwidth]{RefractiveIndexINFO.pdf}
		\label{fig_silicon_dispersion}
	\end{figure}
\end{frame}
%\subsubsection{Multimodal Waveguides}
\begin{frame}{Multimodal Waveguides}
	In silicon waveguides both equations can be verified.\\
	Therefore Eq. \eqref{eq_momentum} becomes:
	\begin{equation}
		\Delta k = 2\,\beta_P(\omega_P) -\beta_S(\omega_S) -\beta_I(\omega_I)
		\label{eq_beta}
	\end{equation}
	\begin{equation*}
		\mathrm{where} \qquad	\beta(\omega) = k_0 n_{eff}(\omega)
	\end{equation*}
	\vspace{-5pt}
	\begin{itemize}
	\item	If the waveguide supports only one mode, phase-matching is only near the frequency of the pump and $\Omega \ll \omega_P$
	\item	If the waveguide is multimodal, there are more matching conditions and also $\Omega \lesssim \omega_P$.
	\end{itemize}
\end{frame}

%\subsection{Phase-matching tuning}
\begin{frame}{Phase-matching tuning}
	\begin{wrapfigure}{r}{0.55\textwidth}
%		\vspace{-10pt}
		\centering
	    \includegraphics[width=.6\textwidth]{narrow_bandwidth.png}
	\end{wrapfigure}
	Phase-matching conditions have narrow bandwidth.

	\vspace{20pt}
	It is difficult to obtain phase-matching at the required wavelength.

	\vspace{20pt}
	We aim to tune with temperature the range where phase-matching is verified.
\end{frame}
%\subsubsection{Thermo-Optic Coefficient}
\begin{frame}{Thermo-Optic Coefficient}
	Refractive index of silicon is influenced from the temperature.

	\begin{equation}
	\begin{split}
		n 	&= n(\omega,T(x,y)) \\
			&= n(\omega,T_{amb}) + \mathrm{TOC}\cdot(T(x,y)-T_{amb})
		\label{eq_refractive_index}
	\end{split}
	\end{equation}

	\vspace{10pt}
	Different modes react differently at the change of temperature.
	\begin{equation}
		n_{eff} = n_{eff}(\omega,T_{amb}) + \mathrm{TOC}_{eff}\cdot(T_H-T_{amb})
		\label{eq_effective_index}
	\end{equation}
\end{frame}
\begin{frame}
	\begin{figure}
		\centering
		\vspace{-10pt}
		\includegraphics[width=.8\textwidth]{effectiveTOC.pdf}
		\label{fig_TOC_effective}
		\vspace{-10pt}
	\end{figure}
	It should be possible to control phase-matching with temperature.
\end{frame}
\section{Computational model}
\begin{frame}{Computational model}
	Definition of the problem: geometry, materials.\\
	Simulation with FEM in Comsol.
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=\textwidth]{geometryR.pdf}
	\end{figure}
\end{frame}
%\subsection{Temperature simulation}
\begin{frame}{Temperature simulation}
	Stationary solution of heat transfer equation:\\
	parameter $T_H$ and boundary conditions.
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=\textwidth]{thermal.pdf}
	\end{figure}
\end{frame}
%\subsection{Mode simulation}
\begin{frame}{Mode simulation}
	Modes as solutions of the Helmholtz equation:\\
	frequency as parameter and boundary conditions.
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=\textwidth]{output_modeTE.pdf}
	\end{figure}
\end{frame}
\begin{frame}{Mode simulation}
	Both TE, TM, and \textit{leaky} modes were obtained.
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=\textwidth]{output_modeTM.pdf}
	\end{figure}
\end{frame}
\section{Data analysis}
\begin{frame}{Data analysis}
	\begin{itemize}
	\item	Data extraction from comsol to matlab.
	\item	Data classification (TE/TM, order).
	\item	Data processing:\\
			\begin{itemize}
			\item	polynomial fit of $\lambda$ and $T_H$.
			\item	evaluation of phase-mismatch \eqref{eq_momentum} with \eqref{eq_beta} and \eqref{eq_effective_index}
					\begin{equation}
					\Delta k = 2	 \frac{2\pi}{ \lambda_P } n_{eff}(\lambda_P,T_H)
								-\frac{2\pi}{ \lambda_S } n_{eff}(\lambda_S,T_H)
								-\frac{2\pi}{ \lambda_I } n_{eff}(\lambda_I,T_H)
					\end{equation}
					and coherence length $L_{coh} = \frac{2\pi}{|\Delta k|}$.
			\end{itemize}
	\item	Data selection: verify condition $L_{coh}/L_{sam} \geq 1$.
	\item	Data aggregation.
	\end{itemize}
\end{frame}
\section{Results}
\begin{frame}{Results: symmetric configuration}
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[height=.8\textheight]{ppaat22.png}
	\end{figure}
\end{frame}
\begin{frame}
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[height=.9\textheight]{baat2.png}
	\end{figure}
\end{frame}
\begin{frame}
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[height=.9\textheight]{ppc2.png}
	\end{figure}
\end{frame}
\begin{frame}
	Combination \texttt{TE 517}, symmetric configuration.
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=.8\textwidth]{bandwidth_movement_TE517_sym.png}
	\end{figure}
\end{frame}
\begin{frame}{Asymmetric configuration}
	We aimed to increase the difference of behaviour between low and high orders.
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[width=\textwidth]{geometryAR.pdf}
	\end{figure}
\end{frame}
\begin{frame}
	\begin{figure}
%		\vspace{-15pt}
		\centering
		\includegraphics[height=.85\textheight]{asym_ppc2.png}
	\end{figure}
\end{frame}
\section{Conclusion}
%\subsection{Improvements}
\begin{frame}{Improvements}
	\begin{itemize}
		\item	The heater should be simulated as a metal conductor media, instead of a simple boundary condition (TM modes).
		\item	Thermo-Optic Coefficient also depends on the wavelength of the signal. It is important for high $\Omega$ conversions.
		\item	Different core geometries should be studied.
		\item	Develop the correlation also between temperature and efficiency of energy conversion.
	\end{itemize}
\end{frame}
%\subsection{Summary}
\begin{frame}{Summary}
	\begin{itemize}
		\item Configurations with a high difference in TOC between their orders are to be preferred.
		\item %Limited interaction between the temperature of material and different modes is a constraint.
		Optimization of the overlap between the thermal and optical spatial profiles in order to maximize the difference of effective TOC in the selected modes.
		\item Thermal tuning of the FWM phase-match is, in fact, possible.
	\end{itemize}
\end{frame}
\end{document}